Polytope notes

### 2020-01-24 notes/polytope-notes.tex
started changes discussed with Raf

  \begin{proposition} \label{prop:halfspace-opt}
    Let the polytopes $\T^r = \{T^r_y \subseteq \reals^d : y \in \Y\}$ and $C$ be given, with $C = \{p: Bp \geq \vec 0\}$ for $B \in \reals^{k \times n}$.
    Then $D(\T^r) \subseteq C$ if and only if there exist $v_1,\dots,v_k \in \reals^d$ with $\inprod{v_i}{x} \leq B_{iy}$ for all $i \in [k], y \in \Y, x \in T^r_y$.
  \end{proposition}
  \begin{proof}
    ($\implies$)
    We are supposing that if $Bp \geq \vec{0}$, then $\vec{0} \in \oplus_p \T^r$.
    In particular, for all vectors $v \in \reals^d$, $0 \leq \sum_{y \in \Y} p(y) \max_{x \in T^r_y} \inprod{v}{x}$.
    \bo{Not sure how to argue from here, unless we get into E matrix stuff. Raf?}

    ($\impliedby$)
    Let $p \in D(\T^r)$; we show $p \in C$.
    By definition, we have $\vec{0} \in \oplus_p \T^r$.
    This implies in particular that for each of the vectors $v_i$,
    \begin{align*}
      0 &\leq \max_{x \in \oplus_p \T^r} \inprod{v}{x}  \\
        &=    \sum_{y \in \Y} p(y) \max_{x \in T^r_y} \inprod{v_i}{x}  \\
        &\leq \sum_{y \in \Y} p(y) B_{iy} .
    \end{align*}
    This proves $Bp \geq 0$, so $p \in C$.
  \end{proof}

  \begin{proposition} \label{prop:vertex-opt}
    Let the polytopes $\T^r = \{T^r_y \subseteq \reals^d : y \in \Y\}$ and $C$ be given, with $C = \conv(\{p^1,\ldots,p^{\ell}\})$.
    Then $D(\T^r) \supseteq C$ if and only if $\vec 0 \in \oplus_{p^j} \T^r$ for all $j \in [\ell]$.
  \end{proposition}
  \begin{proof}
    ($\implies$)
    Because each $p^j \in C$, we have $p^j \in D(\T^r)$, and by definition, this gives $\vec{0} \in \oplus_{p^j} \T^r$.

    ($\impliedby$)
    Let $p \in C$; we show $p \in D(\T^r)$.
    For each $j$, there exist $\{X^j_y \in T^r_y : y \in \Y\}$ with $\vec{0} = \sum_{y \in \Y} p^j(y) X^j_y$.
    Write $p$ as a convex combination of $p^1,\ldots,p^{\ell}$.
    Let $x_y$ be the corresponding convex combination of $X^1_y,\ldots,X^{\ell}_y$.
    Because $X^1_y,\ldots,X^{\ell}_y \in T^r_y$, a convex set, so is $x_y$.
    And $\vec{0} = \sum_{y \in \Y} p(y) x_y$, so $p \in D(\T^r)$.
  \end{proof}

  \begin{corollary} \label{cor:vertex-halfspace-opt}
    Let the polytopes $\T^r = \{T^r_y \subseteq \reals^d : y \in \Y\}$ and $C$ be given, with $C = \conv(\{p^1,\ldots,p^{\ell}\}) = \{p: Bp \geq \vec 0\}$ for $B \in \reals^{k \times n}$.
    We have $D(\T^r) = C$ if and only if both of the following hold:
    \begin{enumerate}
      \item (Halfspace condition) There exist $v_1,\ldots,v_k \in \reals^d$ with $\inprod{v_i}{x} \leq B_{iy}$ for all $i \in [k], y \in \Y, x \in T^r_y$.
      \item (Vertex condition) For all $j \in [\ell]$, $\vec{0} \in \oplus_{p^j} \T^r$.
    \end{enumerate}
  \end{corollary}

@@

  \begin{theorem}
  	Given a convex polytope $C \subseteq \simplex$, there exist polytopes $\T^r$ in dimension $d$ such that $D(\T^r) = C$ if and only if there is a feasible solution to the above quadratic program with parameter $d$.
  \end{theorem}
  \begin{proof}
    By Corollary \ref{cor:vertex-halfspace-opt}, it suffices to show that $\T^r$ satisfying the halfspace and vertex conditions exist if and only if the program is feasible.

>>>>

  \begin{condition}[Halfspace condition]\label{cond:H-condition}
  	A collection of polytopes $\T^r$ and a polytope $C$ in the simplex defined by $C = \{p \in \simplex : Bp \geq \vec 0\}$ \emph{satisfy the halfspace condition} if there exist $v_1, \ldots, v_k \in \reals^d$ such that, for all $i \in [k]$ and $y \in \Y$, for all $x \in T^r_y$, we have $\inprod{v_i}{x} \leq B_{iy}$.
  \end{condition}
  \begin{condition}[Vertex condition]\label{cond:V-condition}
  	A collection of polytopes $\T^r$ and a polytope $C$ in the simplex defined by $C = \conv(\{p^1, \ldots, p^\ell\})$ \emph{satisfy the vertex condition} if for all $j \in [\ell]$, $0 \in \oplus_{p^j} \T^r$. %% Equivalent: there exist $\{x_{jy} \in T_y : j \in [\ell]\}$ such that, for all $j \in [\ell]$, $\sum_{y \in \Y} p^j_y x_{jy} = \vec 0$.
  \end{condition}

  Useful lemmas:
  \begin{lemma} \label{lemma:D-polytope}
    For any $\T^r$, $D(\T^r)$ is a polytope (in particular, is convex).
  \end{lemma}
  \begin{proof}
    \bo{TODO}
  \end{proof}
  \begin{lemma}  \label{lemma:minkowski-support-all}
    The condition $\vec 0 \in \oplus_p \T^r$ holds if and only if, for all $w \in \reals^d$, $\sum_{y \in \Y} p(y) \max_{x \in T^r_y} \inprod{w}{x} \geq 0$.
  \end{lemma}
  \begin{proof}
    Recall by definition, the notation $\oplus_p \T^r = \{\sum_y p(y) x^r_y : x^r_y \in T^r_y (\forall y)\}$.
    In particular, by the halfspace representation of this convex set, $x \in \oplus_p \T^r$ if and only if, for all $w$, $\inprod{w}{x} \leq \max_{x' \in \oplus_p \T^r} \inprod{w}{x'}$.
    Now, $\max_{x' \in \oplus_p \T^r} \inprod{w}{x'} = \sum_y p(y) \max_{x' \in T^r_y} \inprod{w}{x'}$.
    Plugging in $x = \vec{0}$ completes the proof.
  \end{proof}
  \begin{lemma} \label{lemma:minkowski-support-finite}
    Given $\T^r$, there exists a finite collection $w_1,\ldots,w_K \in \reals^d$ for some $K$ such that $\vec 0 \in \oplus_p \T^r$ if and only if, for all $i \in [K]$, $\sum_{y \in \Y} p(y) \max_{x \in T^r_y} \inprod{w_i}{x} \geq 0$.
  \end{lemma}
  \begin{proof}
    [Bo: TODO! The point: If we fix the support of $p$, there are only a finite number of normals that matter, regardless of which $p$ we pick. Varying over all finitely many choices of the support still gives a finite number of normals.]
  \end{proof}
  \begin{lemma} \label{lemma:E-to-B}
    Let $C = \{p : Bp \geq \vec 0 \}$ where $B$ has the minimum possible number of rows to capture $C$, and suppose $C = \{p : Ep \geq \vec 0 \}$.
    Then for each row in $B$ there is some (unique) row in $E$ that is equal to $\alpha B$ for some positive $\alpha$.
  \end{lemma}
  \begin{proof}
    [Bo: ? Should cite]
  \end{proof}

  \begin{theorem} \label{thm:vertex-halfspace-opt}
    Let the polytopes $\T^r = \{T^r_y \subseteq \reals^d : y \in \Y\}$ and $C$ be given, with $C = \conv(\{p^1,\ldots,p^{\ell}\}) = \{p: Bp \geq \vec 0\}$ for $B \in \reals^{k \times n}$.
    We have $D(\T^r) = C$ if and only if both the halfspace and vertex conditions hold.
  \end{theorem}
  \begin{proof}
    ($\implies$)
    Suppose $D(\T^r) = C$.
    First, we note that the vertex condition is immediate: For all $j \in [\ell]$, $p^j \in C$ which gives $p^j \in D(\T^r)$.
    To show the halfspace condition is satisfied, we first construct a matrix $E$ such that $Ep \geq 0 \iff Bp \geq 0$, then use our construction to pick out the necessary vectors $v_1,\dots,v_k$.

    By Lemma \ref{lemma:minkowski-support-finite}, there is a finite collection of vectors $w_1,\dots,w_{K} \in \reals^d$ where $\vec 0 \in \oplus_p \T^r$ if and only if, for all $w_i$, $\sum_y p(y) \max_{x \in T^r_y} \inprod{w_i}{x} \geq 0$.
    Hence, each vector $w_i$ generates a row of a matrix $E \in \reals^{K \times n}$ with $E_{iy} = \max_{x \in T^r_y} \inprod{w_i}{x}$, and we have $0 \in \oplus_p \T^r \iff Ep \geq 0$.
    By assumption of $\Opt(\T^r,C)$, then, we have $Ep \geq 0 \iff Bp \geq 0$.
    By Lemma \ref{lemma:E-to-B}, because $B$ has the minimum possible number of rows, each row of $B$ appears (scaled by some positive constant) as a different row of $E$. Taking the collection of $w_i$ corresponding to these rows and rescaling them by that positive constant, we get a collection of $k$ vectors $v_1,\ldots,v_k \in \reals^d$ with $\max_{x \in T^r_y} \inprod{v_i}{x} = B_{iy}$, hence the halfspace condition is satisfied.

    ($\impliedby$)
    Suppose Conditions \ref{cond:H-condition} and \ref{cond:V-condition} hold.
    Then by the vertex condition, $p^j \in D(\T^r)$ for all $j \in [\ell]$.
    Because $D(\T^r)$ is convex (Lemma \ref{lemma:D-polytope}), this implies $C \subseteq D(\T^r)$.
    To show $D(\T^r) \subseteq C$, let $p \in D(\T^r)$; by definition, $0 \in \oplus_p \T^r$.
    Then in particular for each vector $v_1,\ldots,v_k$ guaranteed by the halfspace condition, we have
    \begin{align*}
      0 &\leq \max_{x \in \oplus_p \T^r} \inprod{v}{x}  \\
        &=    \sum_{y \in \Y} p(y) \max_{x \in T^r_y} \inprod{v_i}{x}  \\
        &\leq \sum_{y \in \Y} p(y) B_{iy} .
    \end{align*}
    This proves $Bp \geq 0$, so $p \in C$.
  \end{proof}

@@

  \begin{theorem}
  	Given a convex polytope $C \subseteq \simplex$, there exist polytopes $\T^r$ in dimension $d$ such that $D(\T^r) = C$ if and only if there is a feasible solution to the above quadratic program with parameter $d$.
  \end{theorem}
  \begin{proof}
    By Theorem \ref{thm:vertex-halfspace-opt}, it suffices to show that $\T^r$ satisfying the halfspace and vertex conditions exist if and only if the program is feasible.
